---
title: "Assignment 4 EDA"
author: "Derek Shat"
date: "11/17/2021"
output: 
  html_document:
    theme: "darkly"
    toc: true
    toc_float:  
      collapsed: false
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r library_setup, message=FALSE}
library(tidyverse) 
library(tidymodels)
library(knitr)
library(GGally)
library(doParallel)
library(ISLR)
data(Smarket)
```

# Data Description
The dataset shows daily data about the S&P 500 stock index between 2001 and 2005. The variables in the data named Lag1 to Lag5 represent the percentage return for the past N days, where N is the number after 'Lag', so Lag5 represents the percentage return for the past 5 days. The Volume variable represents the number of daily shares traded in billions, and direction represents a positive or negative return for the market for the day.

# Plot Pairs
```{r EDA, message = FALSE, fig.width=12, fig.height=8}
market_dt <- Smarket %>%
  as_tibble() %>%
  select(-Today)

ggpairs(market_dt, 
        lower =  list(continuous = wrap("smooth_loess",
                                        color='skyblue'), 
                      combo = "facethist", 
                      discrete = "facetbar", 
                      na = "na"),
        upper = list(continuous = wrap("cor", size=4), 
                      combo = "facethist", 
                      discrete = "facetbar", 
                      na = "na")) +
  theme_minimal() + 
  labs(title = 'Exploratory Data Analysis',
       subtitle = '2001-2005 S&P 500 data')

```

There are no significant correlations between any variables other than between year and volume.

# Descriptive Stats
```{r descriptive_stats}
market_dt %>%
  select(-Year, -Direction) %>% 
  rownames_to_column('id') %>%
  pivot_longer(-id) %>%
  group_by(name) %>%
  summarise(min = min(value), 
            q25 = quantile(value,0.25),
            med = median(value),
            avg = mean(value),
            q75 = quantile(value, 0.75),
            max = max(value),
            IQR = IQR(value),
            sd = sd(value)) %>%
  kable(digits=3)

market_dt %>% count(Direction)
```
